---
title: "Analysing DemoS runs using VedaR"
author: ""
date: "30/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      warning = F,
                      message = F)
```


# {.tabset}

```{r}
#Load packages
library(vedar)
library(tidyverse) #library for data processing and plots
library(plotly) #library for interactive plots
library(igraph) #library for working with network (graph) objects
library(DT) #library for javascript table display
library(kableExtra) #library for html table display
```

```{r variables}
#define variables
filename_base_001 <- "./data/demos_001"
filename_base_007 <- "./data/demos_007"
all_costs_001_file <- "./data/all_costs_001.csv"

```

## Importing the data

* R is efficient at handling data in "long" format: each column is a variable and repeated measures are in separate rows (Excel tables are often use "wide" format)
* The `prep_data()` function imports and combines the data contained in the vd, vds, and vde file. This generates a single long R dataframe (tibble). All strings are in lower case, missing data is replaced with NA (or "annual" for timeslice).
* Here, we import all the data from demos_001 to create a dataframe object `demos_001`, and display the data in an interactive table `demos_001`

```{r import_data}

#import the demos_001 data into a tibble
demos_001 <- prep_data(filename_base = filename_base_001)

#display the first 20 rows of the dataframe
datatable(demos_001)



```

## Comparing the data to veda tables

* We can compare results produced by VedaR to those produced from tables generated by the Veda graphical-user interface. 

* Here, we compare the output of the "All costs" Veda table to comparable output from VedaR.


```{r compare_to_veda}
# load results tables that were generated by Veda DemoS_001 example

all_costs_001 <- read.csv(all_costs_001_file)

all_costs_001 %>%
  select(Attribute, 
         Commodity, 
         Process, 
         Period, 
         Region,
         Pv) %>%
  arrange(Attribute, Process, Period) %>%
  knitr::kable(caption = "Veda csv output")  %>%
  kable_styling(latex_options = "striped")


demos_001 %>% 
  #select rows in which attribute contains the string "cost_" 
  # and process contains "coa
  filter(str_detect(attribute, "cost_"), 
         str_detect(process, "coa")) %>%
  select(attribute, 
         commodity, 
         process, 
         period, 
         region,
         pv) %>%
  arrange(attribute, process, period)  %>%
  knitr::kable(caption = "VedaR data") %>%
  kable_styling(latex_options = "striped")
  
  
```  

## Data Visualisation with R

* A powerful and widely-used data visualisation package in R is ggplot
* Here, we create a plot to display compare costs over years broken down by process, from the demos_001 data

```{r}
data_to_plot <- demos_001

data_to_plot  %>%
   #select the cost attributes
  filter(str_detect(attribute, "(cost_)")) %>%
  # period will be plotted on x-axis, so replace period == NA with "none" for display
  mutate(period = replace_na(period, "None")) %>%
  # create the plot
  ggplot(aes(x = as.character(attribute), 
             y = pv, 
             # the colour is specified by the process column
             fill = process)) + 
  geom_col() +
  #axis labels
  xlab("") +
  ylab("Cumulative period cost") + 
  #remove legend title
  scale_fill_discrete("") +
  #rotate x-label
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  # group by period and region in separate facets
  facet_grid(rows = vars(period),
             cols = vars(region),
             scales = "free") +
  #set background theme of plot
  theme_bw()



```


* We can create a plot for a different dataset by simply changing the data that is passed to the plot.
* Here, we plot the demos_007 data by changing the object data_to_plot. All other lines remaing the same

```{r, out.width='1000px'}
demos_007 <- prep_data(filename_base = filename_base_007)
data_to_plot <- demos_007 %>%
  #filter for processes starting with "elc"
  filter(str_detect(process, "^(elc)"))

data_to_plot %>%
  #select the cost attributes
  filter(str_detect(attribute, "(cost_)")) %>%
  # period will be plotted on x-axis, so replace period == NA with "none" for display
  mutate(period = replace_na(period, "None")) %>%
  # create the plot
  ggplot(aes(x = as.character(attribute), 
             y = pv, 
             # the colour is specified by the process column
             fill = process)) + 
  geom_col() +
  #axis labels
  xlab("") +
  ylab("Cumulative period cost") + 
  #remove legend title
  scale_fill_discrete("") +
  #rotate x-label
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  # group by period and region in separate facets
  facet_grid(rows = vars(period),
             cols = vars(region),
             scales = "free") +
  #set background theme of plot
  theme_bw()


```

* The ggplotly can create interactive plots
* Here, we create an point plot of the demos_007, var_act variable for "elc" processes
```{r}
data_to_plot <- demos_007 %>%
  #filter for processes starting with "elc"
  filter(str_detect(process, "^(elc)")) %>%
  # to get the annual var_act, we sum over time slices - more complex manipulations can be done
  group_by(period, process, region, attribute) %>%
  summarise(pv = sum(pv)) %>%
  ungroup()
 

var_act_plot <- data_to_plot %>%
  #select the "var_act" attribute
  filter(attribute == "var_act") %>%
  # create the plot
  ggplot(
    # make the x-axis a date object
    aes(x = as.Date(paste(period, "01/01", sep = "/")),
             y = pv, 
             colour = process
             )) + 
  geom_point() +
  geom_line() +
  #axis labels
  xlab("") +
  ylab("Annual process activity in period") + 
  #rotate x-label
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  #set background theme of plot
  theme_bw() 

ggplotly(var_act_plot)

```

## Visualising the energy system

* [Sankey diagrams](https://en.wikipedia.org/wiki/Sankey_diagram) can be used to visualise flows between elements in a connected syste,. 
* The VedaR function `make_res()` visualises the system in of the TIMES solution as a Sankey diagram.
* In the current version, the magnitude of flows is  not represented, and all the widths are set to a constant value
```{r echo = T, fig.width=7, , fig.height=5}


res_all <- make_res(dat = demos_007,
                    region_select = "reg1",
                    period_select = 2020,                 
                    node_labels = process_description, 
                    edge_labels = commodity_description, 
                    sankey_width = 1000,
                    sankey_height = 1000,
                    font_size = 10)

res_all



```




## Representing the energy system as a network

* ["Graphs"](https://en.wikipedia.org/wiki/Graph_(discrete_mathematics)) are formal representations of connected networks. Graphs are comprised of nodes and edges.  Nodes and edges can be described by sets of attributes, e.g. Edges can have weights. 
* R has tools for efficient analysis of graph structures. The `igraph` package is one of these [https://igraph.org/r/html/latest/](https://igraph.org/r/html/latest/).
* The VedaR function `make_graph_from_veda_df()` transforms the veda data tibble that was produced by `prep_data()` into an igraph object. If the data contains only a single year, edge weights are assigned based on `var_fin` and `var_fout` attributes.

* Here, we convert the demos_007 data for 2020 and reg1 to an igraph object.

```{r, echo = T}
g <- demos_007 %>%
      filter(period == 2020, 
             region == "reg1") %>%
      make_graph_from_veda_df(node_labels = process,
                              edge_labels = commodity
                              )

# print edge weights
print("Edge weights")
E(g)$weight

```

### Identifying paths

* We can compute all simple paths from a given node using the igraph function `all_simple_paths()` (or between two nodes if the `to = ` argument is specified)

* Here, we create an object `all_mincoa1_paths` that is the list of all paths starting from the mincoa1 process.


```{r, echo = T}

all_mincoa1_paths <- all_simple_paths(g, from = "mincoa1")
all_mincoa1_paths

```

### Check if a string or element appears in a set of paths

* The `vedar` function `check_in_paths()` provides a means of checking whether a
string expression is included in the set of paths. The string needs to  be passed as a (regular expression)[https://www.petefreitag.com/cheatsheets/regex/] 

* Here we check if the string "exp" appears in the set of paths "all_mincoa1_paths"

```{r, echo = T}
check_in_path("(exp)", all_mincoa1_paths)
```

* More complex queries can be created using `igraph` functions. Here, we check whether an "exp" process is included in the paths that are linked to the "coa" commodity as follows

```{r, echo = T}

#find all coa edges
coa_edges <- which(E(g)$commodity == "coa")

#find the start vertices of the coa commodity edges
coa_start_vertices <- ends(g, coa_edges)[,1]

#find all the paths that are linked to the coa_start_vertices

all_coa_paths <- all_simple_paths(g, from = unique(coa_start_vertices))

#check if the string "exp" appears in the set of paths that project from coa_start_vertices
print("Check if 'exp' is in all_coa_paths")
check_in_path("(exp)", all_coa_paths)

#check if the string "dist" appears in the set of paths that project from coa_start_vertices
print("Check if 'dist' is in all_coa_paths")
check_in_path("(dist)", all_coa_paths)



```